---
title: "JP Morgan Healthcare Conference"
output:
  pdf_document: default
  html_notebook: default
---

# Introduction
Motivation: The 2019 JP Morgan Healthcare Conference draws to a close and it got me wondering how much of an impact this conference has on healthcare stock value fluctuations. It's well known that stocks change in response to the speeches, deals, and data presented at the meeting. I've found some sources that attempt to quantify the impact of the meeting on stocks (see below). But I wanted to (a) ask the specific questions below, (b) work with financial data for the first time in R, and (c) practice using tidy data approaches. 

I hypothesize that:   
1. Healthcare stock prices tend to fluctuate more during JPM week than any other week in the same year.  
2. If (1.) is true, then this effect is increasing in magnitude over time. 

Example quantification of impact of JPM week on biotech stocks:  
[https://www.cnbc.com/2017/01/04/betting-on-biotech-during-jpmorgans-big-health-care-conference-pays-off-history-shows.html]  
Quotes:  
*"Biotech has historically outperformed the broader market during The J.P. Morgan Healthcare Conference"  
*"In the past 16 years, the NYSE Arca Biotechnology index (BTK), which measures the performance of 30 biotechnology firms, has outperformed the S&P 500 index by nearly 3 percent during JPMorgan's conference"  


Let's get into the analysis.

Load libraries. 
```{r}
library(quantmod)        # Download stock prices
library(tidyverse)       # Data manipulation / visualization
library(BatchGetSymbols) # Download many stock prices
library(lubridate)       # Dates
```

## Data import  
### Data required:  
*Daily stock prices (avg price and range) for lots of biotechs  
*JPM week dates back to inception of meeting  
  
### Get a list of biotech/pharma/healthcare tickers  
```{r}
# hat tip: following two URLs copied from this Stack Overflow answer: https://stackoverflow.com/questions/25338608/download-all-stock-symbol-list-of-a-market

nasdaq_tickers <- read_csv("http://www.nasdaq.com/screening/companies-by-industry.aspx?exchange=NASDAQ&render=download")
nyse_tickers <- read_csv("http://www.nasdaq.com/screening/companies-by-industry.aspx?exchange=NYSE&render=download")

rbind(nasdaq_tickers, nyse_tickers) %>%
  filter(Sector == "Health Care") %>%
  select(-c(LastSale, MarketCap, `ADR TSO`, `Summary Quote`, X10, IPOyear, Sector)) %>%
  distinct(Name, .keep_all = TRUE)-> healthcare_tickers

head(healthcare_tickers)
```

### Import stock data for these biotechs
```{r}
jpm_start <- as.Date("1983-01-01")
jpm_end   <- as.Date("2017-12-31")

#invisible(capture.output(healthcare_daily_stocks <- BatchGetSymbols(tickers = healthcare_tickers$Symbol, 
#                                           thresh.bad.data = 0.01, 
#                                           bench.ticker = "GSPC", 
#                                           first.date = jpm_start,
#                                           last.date = jpm_end, 
#                                           freq.data = freq.data,
#                                           cache.folder = file.path(tempdir(), 'BGS_Cache')))) # cache in tempdir()
#
# write_csv(healthcare_daily_stocks$df.tickers, path = paste("/Users/katiephd/Box Sync/Katie_Documents/JPM19_stock_analysis/", Sys.Date(), "_healthcare_stock_prices_1983-2017.csv", sep = ""))

healthcare_daily_stocks <- NULL
healthcare_daily_stocks$df.table <- read_csv("/Users/katiephd/Box Sync/Katie_Documents/JPM19_stock_analysis/2019-01-09_healthcare_stock_prices_1983-2017.csv")

```

```{r}
head(healthcare_daily_stocks$df.tickers)
stock_subset <- healthcare_daily_stocks$df.tickers[1:50000, ]
ggplot(stock_subset, aes(x = ref.date, y = price.close)) + 
  geom_line() + 
  facet_wrap(~ticker, scales = 'free_y') 
```

### Import JPM week dates
I couldn't find a website where is listed the JPM weeks over the history of the conference. Based on some searches of the previous names of the conference (Hambrecht & Quist Healthcare Conference, Chase H & Q, now JP Morgan) it seems like it's always early January.   
*First JPM week, Hambrecht & Quist Healthcare Conference, was January 10-12, 1983.  [https://wrhambrecht.com/wp-content/uploads/2015/02/LSFMagWinter-2015-page1.png]  
*Genetech press release puts second JPM week on January 10, 1984.  
*2001: Puts it the week of Jan 12 [https://www.wired.com/2001/01/the-biotech-faves-for-2001/]  
*2018 was January 8 - 11 [https://www.jpmorgan.com/country/US/EN/jpmorgan/investbk/conferences/healthcare]  
  
So I picked likely start dates in each year and then filled in the dates for the rest of the week.   
```{r}
x <- seq(ymd("1983-01-01"),ymd("2018-12-31"),by="1 day")
jpm_start_dates <- x[wday(x,label = TRUE) == "Sun" & 5 <= day(x) & day(x) <= 11  & week(x) <= 2]
jpm_week_dates_by_year <- as.tibble(matrix(ymd("1983-01-01"), nrow = 2018 - 1983 + 1, ncol = 5)) 
colnames(jpm_week_dates_by_year) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
jpm_week_dates_by_year$Monday <- jpm_start_dates
jpm_week_dates_by_year$Tuesday <- jpm_start_dates + days(1) 
jpm_week_dates_by_year$Wednesday <- jpm_start_dates + days(2) 
jpm_week_dates_by_year$Thursday <- jpm_start_dates + days(3) 
jpm_week_dates_by_year$Friday <- jpm_start_dates + days(4) 

jpm_all_dates <- c(jpm_week_dates_by_year$Monday, 
                   jpm_week_dates_by_year$Tuesday,
                   jpm_week_dates_by_year$Wednesday, 
                   jpm_week_dates_by_year$Thursday,
                   jpm_week_dates_by_year$Friday)
```


## Analysis
Cleaning the data, some thoughts:  
*I removed stock prices that are unbelievable high. I found that "The most expensive publicly traded stock of all time is Warren Buffett's Berkshire Hathaway (BRK.A), which is trading at $305,500 per share, as of Oct. 23, 2018))" [https://www.investopedia.com/ask/answers/.../whats-most-expensive-stock-all-time.asp_] so I removed any prices that are higher than this value.  
*I (above) removed companies duplicated between NYSE/NASDAQ. Were these duplications due to companies switching exhanges? Other issues? 
*Other cleaning I could do: check that low prices are always < high prices, check for missing data, etc...   
*A point on missing data: the BatchGetSymbols function does a lot to replace missing data with most recent prices. I could look into the impact of this default parameter on my analysis at a later time.  

```{r}
healthcare_daily_stocks$df.tickers <- healthcare_daily_stocks$df.tickers %>%
  filter(price.high < 305000)
dim(original_healthcare_stocks)
dim(healthcare_daily_stocks$df.tickers)
```
Looks like we got rid of some bad data. Plot some price.high by company over the dates.   
```{r}
high_price_subset <- healthcare_daily_stocks$df.tickers[1:50000, ]
ggplot(high_price_subset, aes(x = ref.date, y = price.high)) + 
  geom_line() + 
  facet_wrap(~ticker, scales = 'free_y') 

```
Looks good. These plots show that BatchGetSymbols() was able to handle companies as the entered and exited the stock market (unlike getSymbols()). 

### Calculate stock price fluctuations
I'm using the word "fluctuations" here because I don't have the technical expertise in finance to use a word like "volatility" which has a couple of specific definitions in finance. By "fluctuations" I mean that I'll calculate the range in price for each stock each week of the year, then I'll compare that range to the range over JPM week. This output will be a relative price range.  

Calculate max net change (range) in price per day.  
```{r}
healthcare_daily_stock_range <- healthcare_daily_stocks$df.tickers %>% 
  filter(ref.date %in% jpm_all_dates) %>%
  filter(price.high < 305000) %>%
  mutate(year = year(ref.date)) %>%
  mutate(price_range = price.high - price.low) %>%
  select(-c(price.open, price.close, volume, price.adjusted, ret.closing.prices, ret.adjusted.prices)) #, price.high, price.low))
  
head(healthcare_daily_stock_range)
summary(healthcare_daily_stock_range$price_range)
```
Calculate max price and min price each JPM week (by year) and list as the price_range.     
```{r}
healthcare_jpm_week_stock_range <- healthcare_daily_stocks$df.tickers %>% 
  filter(ref.date %in% jpm_all_dates) %>%
  filter(price.high < 305000) %>%
  mutate(year = year(ref.date)) %>%
  group_by(ticker, year) %>% 
  summarise(max_price = max(price.high),
            min_price = min(price.low)) %>%
  mutate(price_range = max_price - min_price)
  
head(healthcare_jpm_week_stock_range)
```
Calculate difference in the price range of all subsequent weeks in the year.  
Note, for simplicity here I'm not calculating each week of the year as M-F trading days, I'm using weeks based on date only (not day of week). I'm collecting data for each week starting on January 21st (week >= 3), where weeks are 7 day periods until the end of the year.  
```{r}
healthcare_weekly_stock_range <- healthcare_daily_stocks$df.tickers %>% 
  filter(price.high < 305000) %>%
  mutate(year = year(ref.date), 
         week = week(ref.date)) %>%
  filter(week >= 3) %>%
  group_by(ticker, year, week) %>% 
  summarise(max_price = max(price.high),
            min_price = min(price.low)) %>%
  mutate(price_range = max_price - min_price)

head(healthcare_weekly_stock_range)
```
Compare all other weeks to JPM week.  
I'll start with just 2017 and ticker == ABBV to simplify things, then build back up to all tickers and years. ABBV is "AbbVie is an American publicly traded biopharmaceutical company founded in 2013. It originated as a spin-off of Abbott Laboratories." [https://en.wikipedia.org/wiki/AbbVie_Inc.] 
```{r}
abbv_2017_jpm_week_price_range <- healthcare_jpm_week_stock_range %>%
  filter(year == 2017, ticker == "ABBV")

abbv_2017_weekly_price_ranges <- healthcare_weekly_stock_range %>%
  filter(year == 2017, ticker == "ABBV")

relative_ranges <- abbv_2017_weekly_price_ranges$price_range / abbv_2017_jpm_week_price_range$price_range
relative_ranges
dated_relative_ranges <- cbind(abbv_2017_weekly_price_ranges$week, relative_ranges)

plot(dated_relative_ranges, xlab = "Week", ylab = "Stock price fluctuation: Each week / JPM week", main = "ABBV stock price fluctuations 2017 \n Weekly fluctuations relative to JPMHC week") 
abline(h = 1, col = "red", lty = 2)
```
From this one example, we can see that the JPM week price range difference was not always greater than any other week of the year, but it was in the top 20% of weeks (10 points above red line implies 10 weeks had a price range >1x JPM week price range).    
Are there any years for ABBV for which the JPM week price range was the maximal range for the year?  
```{r}
abbv_jpm_week_price_ranges <- healthcare_jpm_week_stock_range %>%
  filter(ticker == "ABBV")

abbv_weekly_price_ranges <- healthcare_weekly_stock_range %>%
  filter(ticker == "ABBV")

years <- unique(abbv_weekly_price_ranges$year)
jpm_week_range_greater_than_other_week <- matrix(NA, nrow = length(years), ncol = length(unique(abbv_weekly_price_ranges$week)))
row.names(jpm_week_range_greater_than_other_week) <- years
colnames(jpm_week_range_greater_than_other_week) <- unique(abbv_weekly_price_ranges$week)
counter <- 1
for (yr in years){
  abbv_jpm_week_price_ranges_in_year <- abbv_jpm_week_price_ranges %>% filter(year == yr) 
  abbv_weekly_price_ranges_in_year <- abbv_weekly_price_ranges %>% filter(year == yr)
  jpm_week_range_greater_than_other_week[counter, 1:length(abbv_weekly_price_ranges_in_year$price_range)] <- abbv_weekly_price_ranges_in_year$price_range < abbv_jpm_week_price_ranges_in_year$price_range
  counter <- counter + 1
}
jpm_week_range_greater_than_other_week
jpm_week_range_greater_than_other_week_summary <- as.data.frame(rowSums(jpm_week_range_greater_than_other_week, na.rm = TRUE))
jpm_week_range_greater_than_other_week_summary <- cbind(row.names(jpm_week_range_greater_than_other_week_summary), jpm_week_range_greater_than_other_week_summary)
colnames(jpm_week_range_greater_than_other_week_summary) <- c("year", "# weeks with price range smaller than JPM week price range")
ggplot(data = jpm_week_range_greater_than_other_week_summary) + 
  geom_point(aes(y = `# weeks with price range smaller than JPM week price range`, x = year)) + 
  theme_bw()

```
No, not quite. It was nearly the case in 2016.  

Now, let's extend this type of analysis to more tickers and also include a better summary visualization. I'll create a table with years as rows, and then each ticker will be a column with the # weeks price range smaller than JPM week price range. We'll be able to overlay all of these different tickers on the graph and we'll look for trends.  

```{r}
years <- unique(healthcare_jpm_week_stock_range$year)
tickers <- unique(healthcare_weekly_stock_range$ticker)

jpm_week_range_greater_than_other_week <- matrix(NA, nrow = length(years), ncol = length(unique(healthcare_weekly_stock_range$week)))
row.names(jpm_week_range_greater_than_other_week) <- years
colnames(jpm_week_range_greater_than_other_week) <- unique(healthcare_weekly_stock_range$week)
ticker_sums <- NULL

for (tkr in tickers){
  counter <- 1
  ticker_jpm_week_price_ranges <- healthcare_jpm_week_stock_range %>% filter(ticker == tkr)
  ticker_weekly_price_ranges   <- healthcare_weekly_stock_range %>% filter(ticker == tkr)
  current_ticker_count <- jpm_week_range_greater_than_other_week
  for (yr in years){
    jpm_week_price_ranges_in_year <- ticker_jpm_week_price_ranges %>% filter(year == yr) 
    weekly_price_ranges_in_year   <- ticker_weekly_price_ranges %>% filter(year == yr)
    if (length(weekly_price_ranges_in_year$price_range) > 0 & length(jpm_week_price_ranges_in_year$price_range) > 0) {
      current_ticker_count[counter, 1:length(weekly_price_ranges_in_year$price_range)] <- weekly_price_ranges_in_year$price_range < jpm_week_price_ranges_in_year$price_range
    }
    counter <- counter + 1
  }
  ticker_sums <- cbind(ticker_sums, rowSums(current_ticker_count, na.rm = TRUE))
}
row.names(ticker_sums) <- years
ticker_sums <- cbind(years, ticker_sums)
colnames(ticker_sums) <- c("years", tickers)
tkr_sums <- as.data.frame(ticker_sums)
tkr_sums$years <- factor(tkr_sums$years)
```
The data is in "Wide" form right now, which is not very tidy! I'll reshape into "Long" form for ease of compatibility with ggplot()
```{r}
reshaped_ticker_sums <- gather(tkr_sums, ticker, weeks, ABBV:ZYNE, factor_key = TRUE)
View(reshaped_ticker_sums)
```
There are some zero values in this table that should really be NAs, so I'll correct that now.  
```{r}
# Change zero to NA if stock prices don't exist for that year

check_stock_exists_each_year <- healthcare_weekly_stock_range %>% 
  group_by(ticker, year) %>%
  summarise(count = n()) %>% 
  filter(count >= 50) %>%
  mutate(stock_exists = "exists", years = year) %>%
  select(-c(year))

cleaned_ticker_sums <- full_join(reshaped_ticker_sums, check_stock_exists_each_year, by = c("ticker", "years"))  
# Drop NA rows from data
cleaned_ticker_sums$count[is.na(cleaned_ticker_sums$stock_exists)] <- NA
cleaned_ticker_sums <- na.omit(cleaned_ticker_sums)
```

Plot the summary figure.  
```{r}
ggplot(data = cleaned_ticker_sums) + 
  geom_point(aes(x = years, y = weeks, color = factor(ticker)), position = "jitter") + 
  theme(legend.position="none", panel.background = element_rect(fill = "white",
                                colour = "white",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "light grey"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")) + 
  labs(title = "Biotech stocks highly volatile during week of JPM Healthcare Conference", x = "Year", y = "Weeks in year stock price changes less than during JPMHC")
```
Yikes, that's pretty ugly and hard to interpret. Let's try again.  

```{r}
ggplot(data = cleaned_ticker_sums) + 
  geom_point(aes(x = years, y = weeks), position = "jitter") + 
  theme(legend.position="none", panel.background = element_rect(fill = "white",
                                colour = "white",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "light grey"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")) + 
  labs(title = "Biotech stocks highly volatile during week of JPM Healthcare Conference", x = "Year", y = "Weeks in year stock price changes less than during JPMHC")
```


```{r}
ggplot(data = cleaned_ticker_sums) + 
  geom_smooth(aes(x = years, y = weeks), na.rm = TRUE,method = "loess") + 
  theme(legend.position="none", panel.background = element_rect(fill = "white",
                                colour = "white",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "light grey"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")) + 
  labs(title = "Recent increase in biotech stock fluctuations during JPM Healthcare Conference", x = "Year", y = "Weeks in year stock price changes less than during JPMHC")
```
Let's change week per year to a percentage of weeks to ease interpretation.  
```{r}
cleaned_ticker_sums <- cleaned_ticker_sums %>%
  mutate(percent_lower = 100 * weeks/count)
```

Plot percentages.  
```{r}
ggplot(data = cleaned_ticker_sums) + 
  geom_smooth(aes(x = years, y = percent_lower), na.rm = TRUE,method = "loess") + 
  theme(legend.position="none", panel.background = element_rect(fill = "white",
                                colour = "white",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "light grey"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")) + 
  labs(title = "Recent increase in biotech stock fluctuations during JPM Healthcare Conference", x = "Year", y = "% weeks/year stock price fluctuates less than during JPMHC")
```
Bingo! Let's interpret this data in light of my original hypotheses: 

1. Healthcare stock prices tend to fluctuate more during JPM week than any other week in the same year.  
    I'm not yet convinced that this effect is robust. Addition statistical approaches necessary to dig into this. For example, I'm just calculating if the fluctuation is higher or lower (binary output each week) and counting that binary into a number per year. A more nuanced approach would incoporate the magnitude of the fluctuation.  
2. If (1) is true, then this effect is increasing in magnitude over time.  
    This trend seems plausible given the data.  

What confounders complicate the interpretation of the data? 
-Beginning of a fiscal quarter? Control by look at other quarters?
-Every four years January means a new presidential term? Control by comparing years with and without inaugurations? 


## Follow up analyses (for another time):
*Drill down a few individual company examples and track data back to press releases / financial reports / etc...  
*Calculate weekly volatility of each stock. Repeat comparative analyses. 
*Do max volatility weeks coincide with known independent events (elections, terrorist attacks, etc...)? ### Do max volatility weeks coincide with max volatility for other markets? 
*Repeat analysis on other companies (S&P500?) to control for indsutry (healthcare/biotech vs other industries).  
*Perform permutation test for statistical significance of JPM week vs other weeks. 
*Report statistics on trend line. 

## Sources: 
Helpful introductory tutorial to stock data in R by Curtis Miller: https://ntguardian.wordpress.com/2017/03/27/introduction-stock-market-data-r-1/

Example of how GetBatchSymbols works and improves upon quantmod library by Marcelo Perlin: 
https://cran.r-project.org/web/packages/BatchGetSymbols/vignettes/BatchGetSymbols-vignette.html

